CDI:WS Production Bias Report

### 2021-09-21 ws_prod_report.Rmd
add comparison of by-sex CHILDES word freqs to DIF

```
```{r show-new-sex-glimmer,echo= FALSE, warning=FALSE,message=FALSE,error=FALSE, fig.width=3, fig.height=3, fig.cap="GLIMMER plot for sample of items showing most DIF in both directions, as well as a random sample from the middle."}
sex_words <- c(sex_items_to_remove$definition) 
#remove words
d_mat_sex <- d_mat[, ! names(d_mat) %in% sex_words, drop = F]
# there are more IDs in d_mat_sex than in d_demo; remove those
d_mat_sex <- d_mat_sex %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)
#copied from George's work may not be nessesary here
sids = d_mat_sex$data_id
d_mat_sex$data_id = NULL
d_mat_sex = data.matrix(d_mat_sex)
row.names(d_mat_sex) = sids

@@

orig_sexn_n <- get_extreme_item_difficulty_differences(mm_sexn_n, 2.25) # min: -.20 max: 0.74 (mean + 2SD) mean: 0.268
sex_histn_n
```
From seeing this my take away is that there are more female DIF Items total but that the extreme ones are mostly male.
```{r sex-words 3rd, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
nnew_sex_items_to_remove <- orig_sexn_n %>% filter(extreme==T)

@@

>>>>

```{r show-new-sex-glimmer,echo= FALSE, warning=FALSE,message=FALSE,error=FALSE, fig.width=3, fig.height=3, fig.cap="GLIMMER plot for sample of items showing most DIF in both directions, as well as a random sample from the middle."}
sex_words <- c(sex_items_to_remove$definition) 
#remove words
d_mat_sex <- d_mat[, ! names(d_mat) %in% sex_words, drop = F]
# there are more IDs in d_mat_sex than in d_demo; remove those
d_mat_sex <- d_mat_sex %>% filter(is.element(data_id, d_demo$data_id)) %>% arrange(data_id)
#copied from George's work may not be nessesary here
sids = d_mat_sex$data_id
d_mat_sex$data_id = NULL
d_mat_sex = data.matrix(d_mat_sex)
row.names(d_mat_sex) = sids

@@

orig_sexn_n <- get_extreme_item_difficulty_differences(mm_sexn_n, 2.25) # min: -.20 max: 0.74 (mean + 2SD) mean: 0.268
sex_histn_n
```

From seeing this my take away is that there are more female DIF Items total but that the extreme ones are mostly male.

```{r sex-words 3rd, fig.width=3, fig.height=3, echo=FALSE,warning=FALSE,message=FALSE,error=FALSE}
nnew_sex_items_to_remove <- orig_sexn_n %>% filter(extreme==T)

@@

### Suggesting replacement items

```{r}
type_freqs = readRDS("data/type_freqs.rds")
type_counts = readRDS("data/type_counts.rds")
type_counts_smoothed = readRDS("data/type_counts_smoothed.rds")

type_counts_smoothed <- type_counts_smoothed %>%
  pivot_wider(id_cols=uni_lemma, names_from=sex, values_from=count_per_mil) %>%
  mutate(propFemale = female / (female + male),
         FtoM = female / male)

# how biased are the items we suggest removing in the corpus counts?
sex_items_to_remove$definition

tmp <- left_join(orig_sex, type_counts_smoothed, by=c("definition"="uni_lemma"))

tmp %>% #filter(extreme==T) %>%
  ggplot(aes(x=propFemale, y=d_diff, color=extreme)) + 
  geom_point(alpha=.5) + theme_bw() +
  ylab("Female - Male Word Difficulty") +
  xlab("Proportion of Utterances to Females in CHILDES")

subset(tmp, extreme==T & propFemale > .47 & propFemale < .53) # hose & dump

cor.test(tmp$propFemale, tmp$d_diff) # .19, p<.001

# for replacements, want equal-frequency by sex (and high-freq overall) words from CHILDES
normalized_childes_counts = readRDS("data/normalized_childes_counts.rds")

childes_rel_freq <- normalized_childes_counts %>% filter(count_per_mil > 10) %>%
  pivot_wider(id_cols=gloss_stem, names_from=target_child_sex, values_from=count_per_mil) %>%
  mutate(propFemale = female / (female + male))
# 8717 words to choose from
hist(childes_rel_freq$propFemale)
mean(childes_rel_freq$propFemale, na.rm=T) # .497

replacements <- childes_rel_freq %>% 
  filter(propFemale < .51, propFemale > .49) %>%
  filter(!is.element(gloss_stem, en_ws$definition)) # 182
```

1. Of the 22 items with extreme sex DIF, only 2 of those fall very close to being equally often spoken to boys and girls ('hose' and 'dump')
2. There are many items on the CDI with small difficulty difference (DIF) by sex
